---
title: "Biostats - Homework 20"
format: html
editor: visual
---

### Introduction

Maggie Layman 10/22 HW 20

The objective of this assignment is to have you continue practicing the skills you have learned in Chapters 3 and 4 of Getting Started with R and reinforce the ideas we discussed in class regarding the Weissgerber et al. paper. Please complete all of the following steps:

1.  Write the code to set up your workspace.

```{r}
rm(list = ls()) # clears environment
library (tidyverse)
library (here)
```

2.  Read in the dataset called soil_data.csv and name it "soil."

```{r}
soil <- read.csv (here ("data", "soil_data.csv"))
```

3.  Create a histogram for the SOM_percent column in the dataframe. Accept the default number of bins.

```{r}
ggplot(soil,aes(x=SOM_percent))+
  geom_histogram()
```

How would you describe the mean_SOM data based on viewing the histogram? Write your answer below:

The histogram does not show a mean value of SOM, but based on viewing the histogram, I would assume it is close to the mode, which appears to be around 6.

4.  Create a new dataframe called SOM by calculating for each forest the mean soil organic matter and the standard deviation of soil organic matter. Make sure to name variables you create "mean_SOM" and "sd_SOM" so that the next code chunk I give you works.

```{r}
SOM <- soil %>% 
    group_by(Forest) %>%
      summarise(
        mean_SOM = mean (SOM_percent),
        sd_SOM = sd (SOM_percent)
      )
View (SOM)

```

Now summarise Nitrogen

```{r}
Nitrogen <-soil %>%
    group_by(Forest) %>%
      summarise(
        mean_N = mean(Nitrate_N_ppm, na.rm=T),
        sd_N = sd(Nitrate_N_ppm, na.rm=T)
      )
View (Nitrogen)
# returns NA bc Nitrate_N_ppm has NAs
# from now on, we'll use Nitrate_N_ppm, na.rm =TRUE
# if theres a different null value, use which () to solve
  # fix_na <- which (DF,nitrate==-999)
  # DF$nitrate[fix_na] <- NA
```

5.  Now run the following code chunk which will create a bar plot of mean soil organic matter per forest, with standard deviation as error bars.

```{r}
ggplot(SOM, aes(Forest, mean_SOM))+
  geom_errorbar(aes(ymax = mean_SOM + sd_SOM, ymin = mean_SOM - sd_SOM), width = 0.2)+
  geom_col(fill = "white", color = "black")+
  ylab("Mean percent soil organic matter")+
  xlab("Forest name")

```

Based on what you see in the bar chart, how would you describe the pattern of SOM between the 4 different forests? Write your answer here:

Based on the bar chart, the SOM mean and SD of the Kip Tract appear to be much higher than the other forests.

Now create a boxplot of the SOM_percent data. Use Forest as the x-axis variable and overlay the point data on top of the boxplot. Make the points blue and somewhat transparent. Call the x-axis "Forest name" and the y-axis "Percent soil organic matter" (hint: See homework 18...).

```{r}
ggplot (soil, aes (x=Forest,y=SOM_percent))+
  geom_boxplot() + # makes a boxplot
  geom_point(size=2,colour='blue',alpha=0.5)+ # not totally sure what overlay point data at the top means, so I'm just going to increase the size of the points
  xlab("Forest Name") + #xaxis label
  ylab("Percent soil organic matter")+ #yaxis label
  theme_bw()

```

Based on what you see in the box plot, write out your answers the following questions:

-   how would you describe the pattern of SOM between the 4 different forests?

All four forests have the majority of their SOM data clustered \<10. Kip Tract has more spread, while the others have closer data points.

-   What do you notice that is different from how the data were displayed in the bar plot?

This boxplot shows us the spread of the data and any outliers which were absent in the bar chart.

-   Which plot type is more informative? Explain your answer.

The boxplot is more informative because it shows us the frequency of the data collected and is easier to read when looking for patterns in sample size, spread, and outliers.
