---
title: "HW 21"
author: "ML"
format: html
editor: visual
---

Maggie Layman Homework 25 10/25/2024

## Beginning Statistics in R

Working through the first half of Chapter 5 in GSWR

Note that when you read in the ladybird and ozone data sets, you will want to convert a number of variables to factor in order to match the book.

Remember to make notes and commentary outside of code chunks and to comment the code within code chunks.

Set things up

```{r}
# clear environment
rm(list = ls())
# open packages
library(tidyverse)
library(here)
```

import data

```{r}
lady <- read.csv(here("data", "ladybirds.csv"))
glimpse (lady)
```

alter data types

```{r}
lady$Habitat <- as.factor(lady$Habitat)
lady$Site <- as.factor(lady$Site)
lady$colour <- as.factor(lady$colour)
glimpse (lady)
```

## organize the data for plotting

```{r}
totals <- lady %>%
  group_by(Habitat,colour)%>%
    summarise (total.number = sum(number))
View (totals)
```

Make a figure (bar chart) of this data

```{r}
ggplot(totals,aes(x=Habitat,y=total.number, fill=colour))+
  geom_bar(stat='identity', position='dodge')
```

same thing, with the specific colors we want

```{r}
ggplot(totals,aes(x=Habitat,y=total.number, fill=colour))+
  geom_bar(stat='identity', position='dodge')+
    scale_fill_manual(values=c(black="black",red="red" ))
```

analysis: looks like the balck is more common in industrial habitat but can we test this? yes, a chi\^2 test will test that frequencies of color is not equally distributed among the habitats. we will expect to reject the null hypothesis

making x\^2 test

```{r}
# first transform data
lady.mat <- xtabs(number ~ Habitat + colour, data=lady)
lady.mat

# then make test
chisq.test(lady.mat)
```

p is very small, so we can reject the null hypothesis and conclude association btwn Habitat and colour/morph

find more about the chi\^2 test

```{r}
lady.chi <- chisq.test(lady.mat)
names(lady.chi)
```

## two sample t-test

read in new data

```{r}
ozone <- read.csv(here("data", "ozone.csv"))
glimpse (ozone)
```

alter data types

```{r}
ozone$Garden.location <- as.factor(ozone$Garden.location)
ozone$Garden.ID <- as.factor(ozone$Garden.ID)
glimpse(ozone)
```

plot data

```{r}
ggplot(ozone,aes(x=Ozone))+
  geom_histogram(binwidth=10)+
  facet_wrap(~Garden.location,ncol=1) # splits into 2 charts by location
  theme_bw()
```

we might want to test difference in mean ozone levels between 2 groups (location)

now do the t-test

```{r}
t.test(Ozone~Garden.location,data=ozone)
```

another way

```{r}
var.test(Ozone~Garden.location,data=ozone)
```
