---
title: "Homework-25"
author: "ML"
format: html
editor: visual
---

# 11/4 Homework 25

## Working through the code for sections 5.6 and 5.7 in Getting Started with R (pp. 118-129).

## One way ANOVAs

workflow: plot-\>model-\>check assumptions-\>interpret-\>plot again

Getting and Plotting the Data:

looking at water flea growth rates

setup per usual

```{r}
rm (list =ls())
library(ggfortify)
library(tidyverse)
library(here)
```

read in data to use

```{r}
daphnia <- read.csv(here("data", "Daphniagrowth.csv"))
```

plot

```{r}
ggplot(daphnia, 
  aes(x=parasite, y=growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip() #try coordinate flipping to fix the name crowding 
```

substantial variation in the gr among the treatments

construct model

```{r}
model_grow<-lm(growth.rate ~ parasite, data=daphnia)
model_grow
```

check assumptions

```{r}
autoplot(model_grow)
```

q-q plot is questionable, I'll trust the book that it's okay

linear model is appropriate, no concerns for even variance or leverage

test and interpretation

```{r}
anova(model_grow)
```

p is very small, evidencing parasite treatment produces an effect on growth rate

## treatment contrasts

get summary table

```{r}
summary(model_grow)
```

recognize that 1.21391 from the model

it says intercept, but that just represents the 1st alphabetically ordered treatment level (control)

treatment contrasts report differences between the reference level

the numbers associated with each parasite are differences between growth rates with the parasite and the control

use dplyr to get means for each group

```{r}
sumDat<- daphnia%>%
  group_by(parasite)%>%
    summarise(meanGR=mean(growth.rate))
sumDat
```

make a figure

```{r}
ggplot(daphnia, 
  aes(x=parasite, y=growth.rate,colour=factor(parasite)))+
  geom_point(size=5)+
  theme_bw()+
  coord_flip() 
```

second option of getting package

```{r}
install.packages("devtools")
library(devtools)
install_github("sinhrks/ggfortify")
```
