---
title: "Homework-23"
author: "ML"
format: html
editor: visual
---

## Homework 23

Oct 29 2024 - Getting Started with R book (Sections 5.4 - 5.5)

set things up

```{r}
rm(list = ls())
library (tidyverse)
library (here)
```

read & examine plant data

```{r}
plant_gr <- read.csv (here("data", "plant.growth.rate.csv"))
glimpse(plant_gr)
```

make a scatterplot comparing soil moisture to growth

```{r}
ggplot(plant_gr,
  aes(x=soil.moisture.content, y=plant.growth.rate))+
    geom_point()+
    ylab("Plant Growth Rate (mm/week)") +
    theme_bw()
```

make a simple linear regression hypothesizing plant growth as a function of soil moisture content

```{r}
model_pgr <- lm(plant.growth.rate~ soil.moisture.content,data=plant_gr)
model_pgr
```

check the assumptions of the linear model

```{r}
library(ggfortify)
autoplot(model_pgr, smooth.colour=NA)
```

interpret results

```{r}
anova(model_pgr)

```

p is small, reject null hypothesis of no correlation

look at intercept and slope values

```{r}
summary(model_pgr)
```
