---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("Data","CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

    ```{r}
    fcrabs <- crabs %>% 
      filter(Sex == "F")
    ```

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

    predictor - height (ratio)

    response - weight (ratio)

two continuous variables, so run a simple linear regression test

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

first, make a plot

```{r}
f2crabs<-crabs%>% filter(Height<1)

ggplot(f2crabs,
  aes(x=Height, y=Weight))+
    geom_point()+
    theme_bw()
```

positive correlation, seems to have 1 outlier

id estimate slope to be (80-0)/(0.65-0) = 80/0.65 = 123 and change and y -int between -10 and -20

now make a linear regression model

```{r}
model_crabs <- lm(Weight~ Height,data=f2crabs)
model_crabs
```

check for conditions

```{r}
autoplot(model_crabs)
```

res v fit straight line shows a weird bend, but straight line before the outliers, this shows that a linear model is best fit

normal q-q shows normal curve in data

scale location lack of pattern shows evenness in the variance

residuals v leverage shows 2 outliers that have signifcant leverage

```{r}
anova(model_crabs)
```

```{r}
summary(model_crabs)
```

Since the p value is very small, we can reject our null hypothesis and conclude that there is a relationship between height and weight in female crabs (613.6 on 1 and 1223 DF, p-value: \< 2.2e-16).

Null hypothesis: slope is different from 0.

\*\* note from class: Here, i should make another plot with the data and the model that is cleaned up nice
